---
title: "Exercise - 17"
output: html_document
---

### Libraries

``` {r, warning=FALSE, message=FALSE}

# SPDS
library(tidyverse)
library(sf)
library(units)
library(elevatr)

# Data
library(USAboundaries)
library(rnaturalearth)
library(rmapshaper)
library(readxl)

# Visualization
library(gghighlight)
library(ggrepel)
library(knitr)
library(ggthemes)
library(leaflet)
library(leafpop)
library(raster)
library(rgdal)
```

```{r setup, include=FALSE}
cities = readr::read_csv("~/github/geog176A-labs/data/uscities.csv") %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  filter(city == "Goleta") %>%
  st_transform(5070) %>% 
  st_buffer(5000) %>% 
  st_bbox() %>% 
  st_as_sfc() %>% 
  st_as_sf()

r_cities = elevatr::get_elev_raster(cities, z = 11) %>% 
  crop(cities)

writeRaster(r_cities, filename = "data/goleta-elev.tif", overwrite = TRUE)

plot(r_cities)


```

### (Describe properties:)
I'm not sure why but I was unable to see the properties information in the console?  From the "tif" file it appears that black represents sea-level elevation and that all elevation above sea-level is represented in white.  Around the swamp area just south-east of campus it appears that the elevation differs from sea-level to above continuously, creating a grey ring around that area.   

