---
title: "Geography 176A"
author: '[Abigail Porter](https://ucsbaporter.github.io/UCSBaporterW1/first-webpage/index.html)'
subtitle: 'Lab 02: COVID-19 Pandemic'
output:
  html_document:
    theme: journal
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## COVID-19 Data

### Data

```{r, eval = FALSE}
https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv
```

### Start

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(readxl)
library(zoo)

url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'
home = read_csv(url)

read_excel = '../data/PopulationEstimates.xls'

pop = readxl::read_excel("../data/PopulationEstimates.xls", skip = 2) %>% 
  select(pop2019 = POP_ESTIMATE_2019, fips = FIPStxt, State)

CA_pop_2019 = pop %>%
  filter(State == "CA") 

```

```{r, eval = FALSE}

```

## Question 1: Covid-19 California Cases 

```{r, message=TRUE, warning=FALSE}
library(tidyverse)

url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'

home = read_csv(url)

#Step 2

steptwo = home %>% filter(state == "California") %>% group_by(county) %>% 
  mutate(newCase = cases - lag(cases)) %>% arrange(-newCase) %>% ungroup()

#Step 3

#Table 1: 5 counties with most cumulative cases

most_cumulative_cases = steptwo %>% filter(date == max(date)) %>% group_by(county) %>%
  summarize(sum_cases = sum(cases)) %>% arrange(-sum_cases) %>%
  ungroup() %>% slice_max(sum_cases, n = 5)

knitr::kable(most_cumulative_cases, caption = c("Cumulative Case Counts: Top 5 CA counties"),
  col.names = c("County", "Cumulative Cases"))

```



```{r, message=FALSE, warning=FALSE}

#Step 2

steptwo = home %>% filter(state == "California") %>% group_by(county) %>% 
  mutate(newCase = cases - lag(cases)) %>% arrange(-newCase) %>% ungroup()

#Step 3

#Table 2: 5 counties with most new cases
most_new_cases = steptwo %>% filter(date == max(date)) %>% slice_max(newCase, n=5) %>% select (county, newCase)

knitr::kable(most_new_cases, caption = "New Case Counts: Top 5 CA counties", col.names = c("County", "New Cases"))

```


```{r}

pop = readxl::read_excel("../data/PopulationEstimates.xls", skip = 2) %>%
  select(pop2019 = "POP_ESTIMATE_2019", fips = "FIPStxt", state = "State")

CA_pop_covid = right_join(pop, home, by ="fips") %>% 
  filter(date >= max(date, na.rm = TRUE)-13, state.x == "CA") %>% 
  group_by(county) %>%
  mutate(newCase = cases - lag(cases),
         newCase_pc = newCase / pop2019,
         cumulative_pc = cases / pop2019) %>% 
  ungroup()

today = CA_pop_covid %>% 
  filter(date == max(date))

most_per_cap_today  = today %>% 
  slice_max(newCase_pc, n = 5)

most_per_cap_all_time = today %>% 
  slice_max(cumulative_pc, n = 5) %>% 
  select(county = county, cumulative_pc)


#Table 1
pop_most_cumulative_cases = CA_pop_covid %>% 
  filter(date == max(date)) %>% 
  slice_max(cases, n=5) %>% 
  select(county, cumulative_pc)

knitr::kable(pop_most_cumulative_cases, caption = c("Pop Cumulative Case Counts: Top 5 CA counties"),
             col.names = c("County", "Cumulative Cases with Pop (100,000)"))
```

```{r}
#Table 2

pop_most_new_cases = CA_pop_covid %>% filter(date == max(date)) %>% slice_max(newCase_pc, n=5) %>% 
  select(county, newCase_pc) 


knitr::kable(pop_most_new_cases, caption = c("Pop New Case Counts: Top 5 CA counties"),
             col.names = c("County", "New Cases with Pop (100,000)"))
```
```{r, message = FALSE}
#Table 3: Safe counties

pop_safe_county = CA_pop_covid %>% 
  group_by(county) %>% 
  summarize(totNew = sum(newCase, na.rm= TRUE) /  (max(pop2019) /100000)) %>% 
  filter(totNew <= 100)


knitr::kable(pop_safe_county, caption = c("Pop Lowest Case Counts: CA counties"),
             col.names = c("County", "Lowest New Cases with Pop")) 


```
### Results:  Covid-19 cumulative case counts are the highest in the following counties: Los Angeles, Riverside, Orange, San Bernardino, and San Diego.  The highest new cases over the past 14 days are in the following counties: Imperial, Tulare, Merced, Madera, and Yuba.  Despite the covid-19 increases throughout the state the following counties have shown less than 100 new cases over the past 14 days: Alpine, Calaveras, Del Norte, El Dorado, Humboldt, Lake, Mariposa, Modoc, Mono, Nevada, Plumas, Shasta, Siskiyou, Trinity, and Tuolumne.  Based on these results you can see that population density plays a large role in the spread of covid-19 and more remote counties in Northern California have less current spread.
    



## Question 2: Covid-19 New York, California, Louisiana, and Florida

```{r}

Q2 = home %>%
  filter(state %in% c("New York","California", "Louisiana", "Florida")) %>%
  group_by(state, date) %>% summarise(cases = sum(cases)) %>% 
  mutate(newCases = cases - lag(cases),
         roll7 = zoo::rollmean(newCases, 7, fill = NA, align = 'right')) %>% ungroup()  
  
Q2 %>% ggplot(aes(x = date, y = newCases)) + geom_col(aes(y = newCases), col = NA, fill ="#F5B8B5") +
  geom_line(aes(y = roll7), col = "darkred", size = 1) + facet_grid(~state, scale = "free_y") +
  ggthemes::theme_wsj() + theme(legend.position = "right")
  labs(title = paste("Daily Cases in NY, CA, LA, FL")) +
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        plot.title = element_text(size = 10, face = 'bold')) +
  theme(aspect.ratio = .5)
```








```{r}


```
### Results:  (Although I was unable to process the second graph I've studied Covid-19 data since the "Stay in Place" orders were given.)  It is important to review population numbers when reflecting on Covid-19 cases to get a better picture of states that are having large peaks or dips within case numbers.  Without factoring population, states with higher populations will appear to have way more covid-19 cases, but this is not the true picture.  Out of the four states reviewed, California has the largest population overall, but currently does not have the highest covid-19 cases.  Since the first set of graphs do not take population into account California would appear much higher than overall and Louisiana would appear to have low cases.  Since Louisiana has a much lower population than California, Florida, and New York if population was factored in the Louisiana graph would show much higher per population cases.  

